Cost- and Deadline-Constrained Provisioning for Scientific Workflow Ensembles in IaaS Clouds

### 2013-04-09 paper/paper.tex
Re-add sections that were commented-out

% Ewa: the following paragraph can be deleted if necessary
%The paper is organized as follows: after an analysis of related work in
%Section~\ref{sec:related}, Section~\ref{sec:problem} describes the
%infrastructure and application model we are targeting in this paper. In
%Section~\ref{sec:algorithms} we describe the dynamic and static algorithms we
%developed. Section~\ref{sec:performance} presents test scenarios and performance
%metrics, while Section~\ref{sec:results} discusses the results of simulation
%studies of these algorithms. Finally, general conclusions, lessons learned, and
%future work are outlined in Section~\ref{sec:conclusions}.

@@

\underline{Performance Metric}
\label{sec:perf_metric}
In order to precisely define the objective of the algrithms it is necessary to
introduce a metric that can be used to score the performance of the different
algorithms on a given problem (ensemble, budget, and deadline). The simplest
approach is to count the number of workflows in the ensemble that each algorithm
is able to complete within the budget before the deadline, but this metric does
not account for the priority-based utility function specified by the user. Using
the counting approach, a less efficient algorithm may be able to complete a
large number of low-priority workflows by executing the smallest workflows
first. In order to account for the priority, we use an exponential
scoring defined as:
$$ Score(e) = \sum_{w~\in~Completed(e)}{2^{-Priority(w)}} $$
where $Completed(e)$ is the set of workflows in ensemble $e$ that was completed
by the algorithm, and $Priority(w)$ is the priority of workflow $w$ such that
the highest-priority workflow has $Priority(w)=0$, the next highest workflow has
$Priority(w)=1$, and so on. This exponential scoring function gives the highest
priority workflow a score that is higher than all the lower- priority workflows
combined:
$$ 2^{-p} > \sum_{i~=~p+1,~\ldots}2^{-i} $$
%  This scoring is consistent with our definition of the problem, which is to
% complete as many workflows as possible, according to their priorities, given a
% set budget and deadline.

@@

The Workflow-Aware DPDS (WA-DPDS) algorithm extends DPDS by introducing a
workflow admission procedure, which is invoked whenever WA-DPDS sees the first
task of a new workflow at the head of the priority queue (i.e. when no other
tasks from the workflow have been scheduled yet). The admission procedure---
shown in Algorithm~\ref{alg:wa-dpds}---estimates whether there is enough
budget remaining to admit the new workflow; if there is not, then the workflow
is rejected and its tasks are removed from the queue. WA-DPDS compares the
current cost (consumed budget) and remaining budget, taking into account the
cost of currently running VMs, and the cost of workflows that have already
been admitted. In addition, it adds a small safety margin of \$0.10 (10\% of the
compute hour cost) to avoid going over budget. We found the admission procedure useful not only to prevent
low-priority workflows from delaying high-priority ones, but also to reject
large and costly workflows that would overrun the budget and admit smaller
workflows that can efficiently utilize idle resources.

% in 
%ensembles containing workflows of non-uniform sizes. It would also be possible 
%to extend this admission procedure to check other constraints, such as whether
%the estimated critical path of the new workflow exceeds the time remaining
%until the deadline.

@@

%The approach used by SPSS is to plan each workflow in the ensemble in priority
%order, rejecting any workflows that would exceed the dedline or budget. Once the
%plan is complete, the VMs are provisioned and tasks are executed according to
%the schedules given by the plan.

%The disadvantage of the static planning approach used by SPSS is that it is
%sensitive to dynamic changes in the environment and the application that may
%disrupt its carefully constructed plan. For example, if there are provisioning
%delays, or if the runtime estimates for the tasks are inaccurate, then workflow
%execution may diverge from the plan. This issue will be discussed further in
%Sections~\ref{sec:delays} and \ref{sec:variances}.

@@

To plan a workflow, the SPSS algorithm assigns sub-deadlines to each individual
task in the workflow, and then schedules each task so as to minimize the cost of
the task while still meeting its assigned sub-deadline. If each task can be
completed by its deadline in the least expensive way, then the cost of the
entire workflow can be minimized without exceeding the deadline. SPSS assigns
sub-deadlines to each task based on the slack time of the workflow, which is
defined as the amount of extra time that a workflow can extend its critical path
and still be completed by the ensemble deadline. For a workflow $w$, the slack
time of $w$ is: $ ST(w) = d - CP(w) $
where $d$ is the deadline and $CP(w)$ is the critical path of $w$. We
assume that $CP(w) \leq d$, otherwise the workflow cannot be completed by the
deadline and must be rejected. 
%For large ensembles we expect the critical path
%of any individual workflow to be much less than the deadline.

@@

Algorithm~\ref{alg:planworkflow} shows how SPSS creates low-cost plans for each
workflow. The \textsc{PlanWorkflow} procedure first calls
\textsc{DeadlineDistribution} to assign sub-deadlines to tasks.
Then, \textsc{PlanWorkflow} schedules
tasks onto VMs, allocating new VMs when necessary. For each task in the
workflow, the least expensive slot is chosen to schedule the task so that it can
be completed by  its deadline. VMs are allocated in blocks of one billing cycle
(one hour) regardless of the size of the task. When computing the cost of
scheduling a task on a given VM, the algorithm considers idle slots in blocks
that were allocated for previous tasks to be free, while slots in new blocks
cost the full price of a billing cycle. 
For example, if a task has a runtime of
10 minutes, and the price of a block is \$1, then the algorithm will either
schedule the task on an existing VM that has an idle slot larger than 10
minutes for a cost of \$0, or it will allocate a new block on an existing VM,
or provision a new VM, for a cost of \$1. 
If the cost of slots on two different
VMs is equal, then the slot with the earliest start time is chosen. To prevent too
many VMs from being provisioned, the algorithm always prefers to extend the
runtime of existing VMs before allocating new VMs. 
%The result of this is that
%the algorithm will only allocate a new block if there are no idle slots on
%existing blocks large enough or early enough to complete the task by its
%deadline, and it will only allocate a new VM if it cannot add a block to an
%existing VM to complete the task by its deadline.

@@

To evaluate and compare the three proposed algorithms, we developed a cloud
workflow simulator~\cite{CloudWorkflowSimulator} based on CloudSim~\cite{Calheiros2011}. Our
simulation model consists of Cloud, VM and WorkflowEngine entities. The Cloud entity starts and
terminates VM entities using an Amazon EC2-like API. VM entities simulate the
execution of individual tasks, including randomized variations in runtime. The
WorkflowEngine entity manages the scheduling of tasks and the provisioning of 
VMs based on the chosen algorithm. We assume that the VMs have a single core
and execute tasks sequentially. 
%Although CloudSim provides a more advanced infrastructure model which includes
%time-sharing and space-sharing policies, we do not use these features since we
%are interested mainly in the execution of tasks on VMs and high-level workflow
%scheduling and provisioning. 
The simulator reads workflow description files in
a modified version of the DAX format used by the Pegasus Workflow Management
System~\cite{Deelman2005}. 
%The modified file format was created for the
%synthetic workflow generator developed by Bharathi, et al~\cite{Bharathi2008}.

@@

Pareto ensembles contain a small number of larger workflows and a large number
of smaller workflows. Their sizes  
%in a Pareto ensemble 
are chosen according to a Pareto distribution. The distribution was modified so
that the number of large workflows (of size $\geq$ 900) is increased by a small amount to
produce a ``heavy-tail''. This causes Pareto ensembles to have a slightly larger
number of large workflows, which reflects behavior commonly observed in many
computational workloads. 
%An example of the distribution of workflow sizes that
%occurs in a Pareto ensemble is shown in Fig.~\ref{fig:ensemble-pareto}.

%\begin{figure}[ht] 
%    \centering
%    \includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
%    \caption[Histogram of workflow sizes in Pareto ensembles]{Histogram of
%    workflow sizes in Pareto ensembles. Workflow size is measured in number of
%    tasks.}
%    \label{fig:ensemble-pareto}
%\end{figure}

The number of workflows in an ensemble depends on the particular application,
but we assume that ensemble sizes are on the order of between 10 and 100
workflows, typical of the real applications we have examined (see Section~\ref{sec:intro}). 
%For example, the number of
%geographical sites of interest to the users of the CyberShake application in
%the past has been on the order of 100. 
%Second, smaller ensembles consisting of
%just a few workflows can be aggregated into a single workflow, so there is no need to
%treat them as an ensemble. Similarly, when the number of workflows grows and
%each workflow has a large number of tasks, either the deadline and budget
%constraints are low enough to prevent many of the workflows from running, or the
%problem of efficiently allocating them to the resources becomes similar to a
%bag-of-tasks problem, which is easier to solve efficiently.

@@

%The SPSS plan describes what tasks to execute on which VMs, but not when to
%execute them. At runtime, the worklow engine is forced to spend extra money
%to extend the runtime of some VMs so that all of the tasks assigned to that
%VM can be completed. At the same time, other VMs can be terminated early
%because the tasks they were assigned finished earlier than expected. However,
%because of dependencies in the workflow, the latter case is less likely to
%happen, which causes gaps in the schedule. So the net result is that the
%overall cost is increased.

Fig.~\ref{fig:variances-deadline} shows the ratio of simulated makespan to budget.
Interestingly, the deadline constraint is rarely exceeded, even in cases with
very poor quality estimates. For the dynamic algorithms this is a result of the
fact that they can adapt to poor estimates and stop submitting new tasks when
the constraints are reached. Makespan is then computed as the finish time of the
last fully completed workflow from the ensemble. For the static algorithm this
is a result of the way that SPSS schedules workflows, and not of a particularly
clever optimization. The SPSS algorithm tends to schedule workflows early, using
up the budget long before the deadline is reached. This is a consequence of the
deadline distribution function in SPSS, which prevents workflows from starting
late. This is illustrated by the gantt chart in
Fig.~\ref{fig:algorithm-example-spss}, which shows how SPSS tends to pile up
workflows at the beginning of the timeline.
% In comparison, the dynamic algorithms tend to spread out workflow starts over
% the entire duration as shown in Fig.~\ref{fig:algorithm-example-dpds}. The
% consequence of this behavior is that SPSS plans tend to use up the budget but
% leave plenty of time before the deadline.
As a result, when the runtime of the plan is increased by introducing errors,
the SPSS plan has some room to expand without exceeding the deadline.

Note that the algorithms have \textit{not} been changed to account for
inaccurate runtime estimates in this experiment. It is likely that better
performance could be achieved if the algorithms were given a hint as to the
accuracy of the task runtimes. Investigating that optimization is left for
future work. 
%The goal here is to determine how well the current algorithms are
%able to stay within the constraints given inaccurate task runtime estimates.

%It is possible that the static algorithm could be modified to add more
%breathing room in the schedule to account for situations where the runtime
%estimates may be inaccurate. Such a modification would result in slightly
%worse scores when the estimates are good, but would improve scores when the
%estimates are bad. Alternatively, the SPSS algorithm could be modified to
%generate a plan that specifies when to provision and deprovision each VM. In
%that case the workflow engine could be prevented from exceeding the budget
%and deadline constraints, but would prevent some workflows from finishing,
%which may significantly decrease the score. These two modifications are
%subjects for future work.

@@

%As in Fig.~\ref{fig:variances}, the y-axis in each
%plot represents the ratio of the simulated value to the constraint value for
%the budget constraint or the deadline constraint.

@@

% Fig.~\ref{fig:delays_scores} shows the relative performance as a
% percentage of the number of high scores achieved by each algorithm as the
% provisioning delay increases. This figure shows that, as the delay
% increases, the relative performance of the static algorithm increases as
% well. This behavior is a result of the fact that the static algorithm is, in
% essence, cheating by using more time and money than the dynamic algorithms.
% In the previous section we showed how the relative performance of the
% algorithms remains the same when a uniform error is applied to the runtime
% constraints. In that case, the dynamic algorithms adapted to the error by
% rearranging the schedule to accomplish the same amount of work given the
% same deadline and budget. In this case, the dynamic algorithms adapt by
% performing less work (executing fewer workflows) to remain within the
% constraints while accounting for the delays. In both cases, the static
% algorithm completed the same amount of work, but did so by exceeding the
% constraints.

% \begin{figure}[ht]
%   \centering
%   \includegraphics[width=0.5\textwidth]{delays_scores}
%   \caption{Percentage of high scores achieved by each algorithm on Montage 
%   with uniform unsorted ensembles when provisioning delay varies from 0 
%   seconds to 15 minutes.}
%   \label{fig:delays_scores}
% \end{figure}

@@

%This problem differs from previous work on grid and utility grid scheduling
%in that cloud infrastructures provide more control over provisioning, so that
%the number of resources can be adjusted according to the requirements of the
%application. Therefore the problem space becomes larger; it requires not only
%an efficient mapping of tasks to available resources, but also the selection
%of the best resource provisioning plan.

%Formulating the problem as a maximization of the number of prioritized
%workflows completed from the ensemble is also novel and requires workflows to
%be admitted or rejected based on their estimated resource demands. We believe
%that this bi-constrained problem is highly relevant because such constraints
%are commonly imposed on many real-world projects. The approach is also
%directly applicable to grid environments that provide resource reservations
%and charge service units for resource use.

@@

%In addition, we found that SPSS tends to perform better on coarse-grained
%workflows than the dynamic algorithms. For wide, fine-grained workflows, such
%as CyberShake and Montage, however, the dynamic algorithms frequently produce
%performance that is as good or better than SPSS, because they are better able
%to pack the smaller tasks onto idle VMs close to the deadline than SPSS,
%which distributes deadlines to tasks in a way that prevents them from
%starting late.

%For very large workflows and ensembles, we found that the planning time of
%the SPSS algorithm is prohibitive. SPSS often took 10 minutes or more to plan
%ensembles of 100 workflows with 800 tasks each. This suggests that ensembles
%of workflows with tens of thousands of tasks, which are commonly encountered
%in real workflow applications, would take many hours to plan using SPSS. This
%makes the dynamic algorithms much more attractive for large scale problems.

@@

%The results of this study can be applied to develop tools that assist
%researchers in planning their large-scale computational experiments. The
%estimates of cost, runtime, and number of workflows completed that can be
%obtained from both the static algorithms and from the simulation runs,
%constitute valuable hints for planning ensembles and evaluating the
%associated trade-offs.

@@

% This balances the columns by inserting a column break at the given reference
\IEEEtriggeratref{15}

>>>>

The paper is organized as follows: after an analysis of related work in
Section~\ref{sec:related}, Section~\ref{sec:problem} describes the
infrastructure and application model we are targeting in this paper. In
Section~\ref{sec:algorithms} we describe the dynamic and static algorithms we
developed. Section~\ref{sec:performance} presents test scenarios and performance
metrics, while Section~\ref{sec:results} discusses the results of simulation
studies of these algorithms. Finally, general conclusions, lessons learned, and
future work are outlined in Section~\ref{sec:conclusions}.

@@

\underline{Performance Metric}
\label{sec:perf_metric}
In order to precisely define the objective of the algrithms it is necessary to
introduce a metric that can be used to score the performance of the different
algorithms on a given problem (ensemble, budget, and deadline). The simplest
approach is to count the number of workflows in the ensemble that each algorithm
is able to complete within the budget before the deadline, but this metric does
not account for the priority-based utility function specified by the user. Using
the counting approach, a less efficient algorithm may be able to complete a
large number of low-priority workflows by executing the smallest workflows
first. In order to account for the priority, we use an exponential
scoring defined as:
$$ Score(e) = \sum_{w~\in~Completed(e)}{2^{-Priority(w)}} $$
where $Completed(e)$ is the set of workflows in ensemble $e$ that was completed
by the algorithm, and $Priority(w)$ is the priority of workflow $w$ such that
the highest-priority workflow has $Priority(w)=0$, the next highest workflow has
$Priority(w)=1$, and so on. This exponential scoring function gives the highest
priority workflow a score that is higher than all the lower- priority workflows
combined:
$$ 2^{-p} > \sum_{i~=~p+1,~\ldots}2^{-i} $$
This scoring is consistent with our definition of the problem, which is to
complete as many workflows as possible, according to their priorities, given a
set budget and deadline.

@@

The Workflow-Aware DPDS (WA-DPDS) algorithm extends DPDS by introducing a
workflow admission procedure, which is invoked whenever WA-DPDS sees the first
task of a new workflow at the head of the priority queue (i.e. when no other
tasks from the workflow have been scheduled yet). The admission procedure---
shown in Algorithm~\ref{alg:wa-dpds}---estimates whether there is enough
budget remaining to admit the new workflow; if there is not, then the workflow
is rejected and its tasks are removed from the queue. WA-DPDS compares the
current cost (consumed budget) and remaining budget, taking into account the
cost of currently running VMs, and the cost of workflows that have already
been admitted. In addition, it adds a small safety margin of \$0.10 (10\% of the
compute hour cost) to avoid going over budget. We found the admission procedure useful not only to prevent
low-priority workflows from delaying high-priority ones, but also to reject
large and costly workflows that would overrun the budget and admit smaller
workflows that can efficiently utilize idle resources in 
ensembles containing workflows of non-uniform sizes. It would also be possible 
to extend this admission procedure to check other constraints, such as whether
the estimated critical path of the new workflow exceeds the time remaining
until the deadline.

@@

The approach used by SPSS is to plan each workflow in the ensemble in priority
order, rejecting any workflows that would exceed the dedline or budget. Once the
plan is complete, the VMs are provisioned and tasks are executed according to
the schedules given by the plan.

The disadvantage of the static planning approach used by SPSS is that it is
sensitive to dynamic changes in the environment and the application that may
disrupt its carefully constructed plan. For example, if there are provisioning
delays, or if the runtime estimates for the tasks are inaccurate, then workflow
execution may diverge from the plan. This issue will be discussed further in
Sections~\ref{sec:delays} and \ref{sec:variances}.

@@

To plan a workflow, the SPSS algorithm assigns sub-deadlines to each individual
task in the workflow, and then schedules each task so as to minimize the cost of
the task while still meeting its assigned sub-deadline. If each task can be
completed by its deadline in the least expensive way, then the cost of the
entire workflow can be minimized without exceeding the deadline. SPSS assigns
sub-deadlines to each task based on the slack time of the workflow, which is
defined as the amount of extra time that a workflow can extend its critical path
and still be completed by the ensemble deadline. For a workflow $w$, the slack
time of $w$ is: $ ST(w) = d - CP(w) $
where $d$ is the deadline and $CP(w)$ is the critical path of $w$. We
assume that $CP(w) \leq d$, otherwise the workflow cannot be completed by the
deadline and must be rejected. 
For large ensembles we expect the critical path
of any individual workflow to be much less than the deadline.

@@

Algorithm~\ref{alg:planworkflow} shows how SPSS creates low-cost plans for each
workflow. The \textsc{PlanWorkflow} procedure first calls
\textsc{DeadlineDistribution} to assign sub-deadlines to tasks.
Then, \textsc{PlanWorkflow} schedules
tasks onto VMs, allocating new VMs when necessary. For each task in the
workflow, the least expensive slot is chosen to schedule the task so that it can
be completed by  its deadline. VMs are allocated in blocks of one billing cycle
(one hour) regardless of the size of the task. When computing the cost of
scheduling a task on a given VM, the algorithm considers idle slots in blocks
that were allocated for previous tasks to be free, while slots in new blocks
cost the full price of a billing cycle. 
For example, if a task has a runtime of
10 minutes, and the price of a block is \$1, then the algorithm will either
schedule the task on an existing VM that has an idle slot larger than 10
minutes for a cost of \$0, or it will allocate a new block on an existing VM,
or provision a new VM, for a cost of \$1. 
If the cost of slots on two different
VMs is equal, then the slot with the earliest start time is chosen. To prevent too
many VMs from being provisioned, the algorithm always prefers to extend the
runtime of existing VMs before allocating new VMs. The result of this is that
the algorithm will only allocate a new block if there are no idle slots on
existing blocks large enough or early enough to complete the task by its
deadline, and it will only allocate a new VM if it cannot add a block to an
existing VM to complete the task by its deadline.

@@

To evaluate and compare the three proposed algorithms, we developed a cloud
workflow simulator~\cite{CloudWorkflowSimulator} based on CloudSim~\cite{Calheiros2011}. Our
simulation model consists of Cloud, VM and WorkflowEngine entities. The Cloud entity starts and
terminates VM entities using an Amazon EC2-like API. VM entities simulate the
execution of individual tasks, including randomized variations in runtime. The
WorkflowEngine entity manages the scheduling of tasks and the provisioning of 
VMs based on the chosen algorithm. We assume that the VMs have a single core
and execute tasks sequentially. 
Although CloudSim provides a more advanced infrastructure model which includes
time-sharing and space-sharing policies, we do not use these features since we
are interested mainly in the execution of tasks on VMs and high-level workflow
scheduling and provisioning. 
The simulator reads workflow description files in
a modified version of the DAX format used by the Pegasus Workflow Management
System~\cite{Deelman2005}. 
%The modified file format was created for the
%synthetic workflow generator developed by Bharathi, et al~\cite{Bharathi2008}.

@@

Pareto ensembles contain a small number of larger workflows and a large number
of smaller workflows. Their sizes  
%in a Pareto ensemble 
are chosen according to a Pareto distribution. The distribution was modified so
that the number of large workflows (of size $\geq$ 900) is increased by a small amount to
produce a ``heavy-tail''. This causes Pareto ensembles to have a slightly larger
number of large workflows, which reflects behavior commonly observed in many
computational workloads. 
An example of the distribution of workflow sizes that
occurs in a Pareto ensemble is shown in Fig.~\ref{fig:ensemble-pareto}.

\begin{figure}[ht] 
    \centering
    \includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
    \caption[Histogram of workflow sizes in Pareto ensembles]{Histogram of
    workflow sizes in Pareto ensembles. Workflow size is measured in number of
    tasks.}
    \label{fig:ensemble-pareto}
\end{figure}

The number of workflows in an ensemble depends on the particular application,
but we assume that ensemble sizes are on the order of between 10 and 100
workflows, typical of the real applications we have examined (see Section~\ref{sec:intro}). 
For example, the number of
geographical sites of interest to the users of the CyberShake application in
the past has been on the order of 100. 
Second, smaller ensembles consisting of
just a few workflows can be aggregated into a single workflow, so there is no need to
treat them as an ensemble. Similarly, when the number of workflows grows and
each workflow has a large number of tasks, either the deadline and budget
constraints are low enough to prevent many of the workflows from running, or the
problem of efficiently allocating them to the resources becomes similar to a
bag-of-tasks problem, which is easier to solve efficiently.

@@

The SPSS plan describes what tasks to execute on which VMs, but not when to
execute them. At runtime, the worklow engine is forced to spend extra money
to extend the runtime of some VMs so that all of the tasks assigned to that
VM can be completed. At the same time, other VMs can be terminated early
because the tasks they were assigned finished earlier than expected. However,
because of dependencies in the workflow, the latter case is less likely to
happen, which causes gaps in the schedule. So the net result is that the
overall cost is increased.

Fig.~\ref{fig:variances-deadline} shows the ratio of simulated makespan to budget.
Interestingly, the deadline constraint is rarely exceeded, even in cases with
very poor quality estimates. For the dynamic algorithms this is a result of the
fact that they can adapt to poor estimates and stop submitting new tasks when
the constraints are reached. Makespan is then computed as the finish time of the
last fully completed workflow from the ensemble. For the static algorithm this
is a result of the way that SPSS schedules workflows, and not of a particularly
clever optimization. The SPSS algorithm tends to schedule workflows early, using
up the budget long before the deadline is reached. This is a consequence of the
deadline distribution function in SPSS, which prevents workflows from starting
late. This is illustrated by the gantt chart in
Fig.~\ref{fig:algorithm-example-spss}, which shows how SPSS tends to pile up
workflows at the beginning of the timeline.
In comparison, the dynamic algorithms tend to spread out workflow starts over
the entire duration as shown in Fig.~\ref{fig:algorithm-example-dpds}. The
consequence of this behavior is that SPSS plans tend to use up the budget but
leave plenty of time before the deadline.
As a result, when the runtime of the plan is increased by introducing errors,
the SPSS plan has some room to expand without exceeding the deadline.

Note that the algorithms have \textit{not} been changed to account for
inaccurate runtime estimates in this experiment. It is likely that better
performance could be achieved if the algorithms were given a hint as to the
accuracy of the task runtimes. Investigating that optimization is left for
future work. 
The goal here is to determine how well the current algorithms are
able to stay within the constraints given inaccurate task runtime estimates.

It is possible that the static algorithm could be modified to add more
breathing room in the schedule to account for situations where the runtime
estimates may be inaccurate. Such a modification would result in slightly
worse scores when the estimates are good, but would improve scores when the
estimates are bad. Alternatively, the SPSS algorithm could be modified to
generate a plan that specifies when to provision and deprovision each VM. In
that case the workflow engine could be prevented from exceeding the budget
and deadline constraints, but would prevent some workflows from finishing,
which may significantly decrease the score. These two modifications are
subjects for future work.

@@

As in Fig.~\ref{fig:variances}, the y-axis in each
plot represents the ratio of the simulated value to the constraint value for
the budget constraint or the deadline constraint.

@@

Fig.~\ref{fig:delays_scores} shows the relative performance as a
percentage of the number of high scores achieved by each algorithm as the
provisioning delay increases. This figure shows that, as the delay
increases, the relative performance of the static algorithm increases as
well. This behavior is a result of the fact that the static algorithm is, in
essence, cheating by using more time and money than the dynamic algorithms.
In the previous section we showed how the relative performance of the
algorithms remains the same when a uniform error is applied to the runtime
constraints. In that case, the dynamic algorithms adapted to the error by
rearranging the schedule to accomplish the same amount of work given the
same deadline and budget. In this case, the dynamic algorithms adapt by
performing less work (executing fewer workflows) to remain within the
constraints while accounting for the delays. In both cases, the static
algorithm completed the same amount of work, but did so by exceeding the
constraints.

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.5\textwidth]{delays_scores}
  \caption{Percentage of high scores achieved by each algorithm on Montage 
  with uniform unsorted ensembles when provisioning delay varies from 0 
  seconds to 15 minutes.}
  \label{fig:delays_scores}
\end{figure}

@@

This problem differs from previous work on grid and utility grid scheduling
in that cloud infrastructures provide more control over provisioning, so that
the number of resources can be adjusted according to the requirements of the
application. Therefore the problem space becomes larger; it requires not only
an efficient mapping of tasks to available resources, but also the selection
of the best resource provisioning plan.

Formulating the problem as a maximization of the number of prioritized
workflows completed from the ensemble is also novel and requires workflows to
be admitted or rejected based on their estimated resource demands. We believe
that this bi-constrained problem is highly relevant because such constraints
are commonly imposed on many real-world projects. The approach is also
directly applicable to grid environments that provide resource reservations
and charge service units for resource use.

@@

In addition, we found that SPSS tends to perform better on coarse-grained
workflows than the dynamic algorithms. For wide, fine-grained workflows, such
as CyberShake and Montage, however, the dynamic algorithms frequently produce
performance that is as good or better than SPSS, because they are better able
to pack the smaller tasks onto idle VMs close to the deadline than SPSS,
which distributes deadlines to tasks in a way that prevents them from
starting late.

For very large workflows and ensembles, we found that the planning time of
the SPSS algorithm is prohibitive. SPSS often took 10 minutes or more to plan
ensembles of 100 workflows with 800 tasks each. This suggests that ensembles
of workflows with tens of thousands of tasks, which are commonly encountered
in real workflow applications, would take many hours to plan using SPSS. This
makes the dynamic algorithms much more attractive for large scale problems.

@@

The results of this study can be applied to develop tools that assist
researchers in planning their large-scale computational experiments. The
estimates of cost, runtime, and number of workflows completed that can be
obtained from both the static algorithms and from the simulation runs,
constitute valuable hints for planning ensembles and evaluating the
associated trade-offs.

@@

% This balances the columns by inserting a column break at the given reference
%\IEEEtriggeratref{15}
